---
title: "Groupwork2"
author: "Kuan"
date: "2023-03-10"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, message = FALSE, warning = FALSE)
```

```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```

```{r loadpackage,echo=TRUE, eval=TRUE, warning=FALSE,message=FALSE}
library(readr)
library(tidyverse)
library(moderndive)
library(gapminder)
library(sjPlot)
library(stats)
library(jtools)
library(janitor)
library(areaplot)
library(dplyr)
library(skimr)
library(kableExtra)
library(gridExtra)
library(ggplot2)
library(stringr)
library(ISLR)
library(plotly)
library(MASS)
library(broom)
```

# Data pre-processing
## Remove missing value
In the raw dataset, there are some missing data about mean altitude and harvested, so before analysis data we remove missing values.
```{r data,echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
dataset13 <- read_csv("dataset13.csv")
dim(dataset13)
#remove NA
newdataset<- na.omit(dataset13)
dim(newdataset)
```

## data cleaning
```{r data cleaning,fig.cap = "\\label{fig:scat} Boxplots of mean altitude of the growers farm/left, aroma values/right in different qualityclass", fig.pos = "H"}
#boxplot about mean altitude by Qualityclass
boxplot_altitude<-ggplot(data = newdataset, mapping = aes(x = factor(Qualityclass), y = altitude_mean_meters, fill = Qualityclass)) +
  geom_boxplot() +
  labs(x = "Qualityclass", y = "Mean altitude of the growers farm/m", title = "Mean altitude in qualityclass")
#boxplot about Aroma
boxplot_aroma<-ggplot(data = newdataset, mapping = aes(x = factor(Qualityclass), y = aroma, fill = Qualityclass)) +
  geom_boxplot() +
  labs(x = "Qualityclass", y = "aroma values", title = "aroma values in qualityclass")
#remove outliers
grid.arrange(boxplot_altitude,boxplot_aroma,ncol=2)
newdataset<- newdataset%>%
  arrange(desc(altitude_mean_meters))
newdataset<- newdataset[-c(1:4),]
newdataset<- newdataset%>%
  arrange(aroma)
newdataset<- newdataset[-1,]
```

The boxplot about mean altitude shows that there are some outliers. Four of them are more than 10000 metres, obviously they are wrong datas, so we remove them.From the boxplot about Aroma, we find there is a wrong value which equals zero and remove it from the dataset.

# Suitable numerical summaries and visualizations 
```{r summaries1,echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
#summary of numerical explanatory variables
newdata_summary<- newdataset%>%
  dplyr::select(aroma,flavor,acidity,category_two_defects,altitude_mean_meters)
my_skim <- skim_with(numeric = sfl(hist = NULL), 
                    base = sfl(n = length))
my_skim(newdata_summary) %>%
  transmute(Variable=skim_variable, n = n, Mean=numeric.mean, SD=numeric.sd,
            Min=numeric.p0, Median=numeric.p50,  Max=numeric.p100,
            IQR = numeric.p75-numeric.p50) %>%
  kable(caption = '\\label{tab:summaries1} Summary statistics on the different numerical explanatory variables of coffee.', booktabs = TRUE, linesep = "", digits = 2) %>%
  kable_styling(font_size = 10, latex_options = "HOLD_position")
```

Table1 shows that the mean values of Aroma grade, Flavor grade and Acidity grade are both approximately 7.5. There are large differences of category two defects between different coffee beans, as some have no defective product, but some have 47 in the batch of coffee beans tested. Similarly, the difference in mean altitude is distinct.

```{r summaries2,echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
#summary of categorical explanatory variables
#country of origin
data_country<- newdataset %>%
  group_by(country_of_origin) %>%
  summarise(n=n())
data_country
newdataset %>%
  tabyl(country_of_origin, Qualityclass) %>%
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  adorn_ns() %>%
  kable(caption = '\\label{tab1:origin} Summary statistics on country of origin.') %>%
  kable_styling(latex_options = "HOLD_position")
```

The summary table shows that there are total 33 countries in the dataset, and 200 observations are from Mexico, which is the most, but some countries have only one observation. We also note that there are 6 countries like Laos only have poor qualityclass of coffee, the qualityclass of Ethiopia and United States(Hawaii) are all good. There also have 3 countries' qualityclass is half and half.

```{r summaries3,echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
#harvested
data_harvested<- newdataset %>%
  group_by(harvested) %>%
  summarise(n=n())
data_harvested
newdataset %>%
  tabyl(harvested, Qualityclass) %>%
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  adorn_ns() %>%
  kable(caption = '\\label{tab1:harvested} Summary statistics on
harvested.') %>%
  kable_styling(latex_options = "HOLD_position")
```

The summary table shows that the information is collected from 2010 to 2018, and 255 observations is from 2012 which is the most. We also note that in 2010 the propotion of good qualityclass is highest, which is 76.9%. The lowest is 39.2% in 2012.

## Country of origin
```{r plot_country,echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE, out.width='90%',fig.align='center',fig.cap="\\label{fig:plot1}Propotion of Qualityclass by countries of origin.",fig.pos='H'}
ggplot(newdataset, aes(x=Qualityclass, y=..prop.., group=country_of_origin, fill=country_of_origin))+
  geom_bar(position = "dodge", stat="count")+
  labs(y="Proportion")
```

The plot shows the propotion of Good qualityclass vs poor qualityclass between different country of origin. we can see some countries are all good quality, some are all poor quality.So we can fit a logistic regression model to determine whether the qualityclass of coffee can be predicted from their country of origin.

## Aroma, Flavour and Acidity
```{r plot_three_grade,echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE, out.width='90%',fig.align='center', fig.cap = "\\label{fig:plot2} Boxplot of aroma in different qualityclass", fig.pos = "H"}
#boxplot of Aroma grade
ggplot(data = newdataset, mapping = aes(x = factor(Qualityclass), y = aroma, fill = Qualityclass)) +
  geom_boxplot() +
  labs(x = "Qualityclass", y = "Aroma",
       title = "Aroma in different qualityclass")
#boxplot of Flavour grade
ggplot(data = newdataset, mapping = aes(x = factor(Qualityclass), y = flavor, fill = Qualityclass)) +
  geom_boxplot() +
  labs(x = "Qualityclass", y = "Flavor",
       title = "Flavor in different qualityclass")
#boxplot of Acidity grade
ggplot(data = newdataset, mapping = aes(x = factor(Qualityclass), y = acidity, fill = Qualityclass)) +
  geom_boxplot() +
  labs(x = "Qualityclass", y = "Acidity",
       title = "Acidity in different qualityclass")
```

The features of these three quality scores are similar. The boxplots show that coffee with good quality have higher grade in Aroma, Flavour and Acidity than poor. So we can fit a logistic regression model to see whether Aroma, Flavour and Acidity are significant predictors of the odds of qualityclass of coffee beans.

## Count of defects
```{r plot_count_of_defects,echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE, out.width='90%',fig.align='center', fig.cap = "\\label{fig:plot5} Boxplot of Count of category 2 type defects in different qualityclass", fig.pos = "H"}
ggplot(data = newdataset, mapping = aes(x = factor(Qualityclass), y = category_two_defects, fill = Qualityclass)) +
  geom_boxplot() +
  labs(x = "Qualityclass", y = "Count of category 2 type defects",
       title = "Count of category 2 type defects in different qualityclass")
```

The boxplot shows that the poor quality coffee beans have more defective products than good quality ones, and there are more outliers in poor quality coffee beans.So we can fit a logistic regression model to see whether count of category 2 type defects is a significant predictor of the odds of qualityclass of coffee beans.

## Mean altitude
```{r plot_altitude,echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE, out.width='90%',fig.align='center', fig.cap = "\\label{fig:plot6} Boxplot of mean altitude of the growers farm in different qualityclass", fig.pos = "H"}
ggplot(data = newdataset, mapping = aes(x = factor(Qualityclass), y = altitude_mean_meters, fill = Qualityclass)) +
  geom_boxplot() +
  labs(x = "Qualityclass", y = "Mean altitude of the growers farm/m",
       title = "Mean altitude of the growers farm in different qualityclass")
```

The boxplot shows that the mean altitude of good quality coffee beans are higher than poor quality ones. we can notice that the poor quality have more outliers. So we can fit a logistic regression model to see whether mean altitude of the growers is a significant predictor of the odds of qualityclass of coffee beans.

## harvested
```{r plot_harvested,echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE, out.width='90%',fig.align='center',fig.align='center',fig.cap="\\label{fig:plot7}Propotion of Qualityclass by Harvested.",fig.pos='H'}
ggplot(newdataset, aes(x=Qualityclass, y=..prop.., group=harvested, fill=harvested))+
  geom_bar(position = "dodge", stat="count")+
  labs(y="Proportion")
prop<-newdataset %>%
  tabyl(harvested, Qualityclass) %>%
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  adorn_ns()
prop$Good<-str_sub(string=prop$Good, start=1, end=5)
prop$Good<-as.factor(prop$Good)
ggplot(prop, aes(x=harvested, y=Good, group=1))+
  geom_line()+
  labs(y="Proportion")
```

The line plot shows the propotion of Good qualityclass is highest in 2010,which is about 75%. We can fit a logistic regression model to determine whether the qualityclass of coffee can be predicted from harvested years.

## Calculating Correlation
In order to prepare for subsequent improvement and selection of variables during modelling, we firstly calculated the correlation between every two numerical variables.
```{r calculating correlation,echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
newdataset[,2:6]%>%
  cor()%>%
  kable(caption='\\label{tab:correlation} correlation between 5 numerical variables')%>%
  kable_styling(font_size = 10, latex_options = "hold_position")
```

Table \ref{tab:correlation} shows the correlation between every two variables including aroma, flavor, acidity, category_two_defects and altitude_mean_meters. We can see that the correlation between aroma& flavor (0.725) and the correlation between flavor&acidity (0.744) are both more than 0.7, which means these pairs have strong positive correlation. There is also a moderate correlation between aroma&acidity (0.591), while the correlation between other pairs are relatively weak.

## Processing Non-numerical Data
For non-numerical data, including country_of_origin, Qualityclass and harvested(year), we set the country_of_origin and harvested(year) as factors. While as a qualitative variable, we converted Qualityclass into dummy variables, 'poor' to '0' and 'good' to '1'.
```{r processing non-numerical data,echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE,results="hide"}
names(newdataset)
newdataset$country_of_origin<- as.factor(newdataset$country_of_origin)
newdataset$Qualityclass<- ifelse(newdataset$Qualityclass=='Poor',0,1)
newdataset$harvested <- as.factor(newdataset$harvested)
```
\newpage



# Formal Data Analysis
We used GLM to fit a logistic regression model with Qualityclass as the binary response variable, and country_of_origin, aroma, flavor, acidity, category_two_defects, altitude_mean_meaters and harvested as the explanatory variables. A summary of the model and a graph showing the points estimate for the log-odds with their corresponding 95% confidence interval are obtained as results.

## Basic GLM
```{r glm,echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
mod.cafe <- glm(Qualityclass ~ country_of_origin + aroma + flavor + acidity + category_two_defects + 
                  altitude_mean_meters + harvested, data = newdataset, family = binomial(link = "logit"))
print(summary(mod.cafe)$call)
tidy(mod.cafe)
AIC(mod.cafe)
#AIC = 543
```

##Plot of distribution
```{r glmplot, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE,out.width = '100%', fig.align = "center", fig.pos = "H", fig.cap="\\label{fig:glm plot} Odds of various factors influencing the quality of coffee(basic GLM model)"}
plot_model(mod.cafe, show.values = TRUE,
           title = "", show.p = FALSE, value.offset = 0.5)
```
In the results we can see that aroma, flavor and acidity has coefficients of 5.19, 8.56, 5,23 separately, indicating comparatively strong positive influence on cafe quality, whilst category_two_defects and altitude_mean_meters do not appear to have much impact. For country_of_origin and harvested, different countries and vintages have different degrees of influence on the quality of coffee. Varies from Thailand(2.56) to India(-2.99). In addition, only the harvested of 2018 shows a little positive impact on cafe quality(2.03), while other variables do not appear to be strongly influential.
\newpage



## GLM Stepwise
In the previous basic GLM we fitted a model with AIC of 543, wondering whether there is better regression to fit the data after selecting only the influencial variables, we then decided to use stepwise regression to improve our model.
```{r glm stepwise,eval=TRUE, warning=FALSE, message=FALSE,}
# Fit a glm using stepwise regression with AIC as the criterion
model.step <- stepAIC(glm(Qualityclass ~ country_of_origin + aroma + flavor + acidity + category_two_defects + altitude_mean_meters + harvested, data = newdataset, family = binomial(link = "logit")), direction = "both", trace = FALSE)
```

```{r glmstepwiseplot, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE,out.width = '100%', fig.align = "center", fig.pos = "H", fig.cap="\\label{fig:glm stepwise plot} Odds of various factors influencing the quality of coffee(basic GLM model)"}
# Print the selected model
print(summary(model.step)$call)
tidy(model.step)
AIC(model.step)
#AIC = 537
plot_model(model.step, show.values = TRUE,
           title = "", show.p = FALSE, value.offset = 0.50)
```
Using stepwise regression, we fitted a model with AIC of 537, which is relatively smaller than 543 in our first basic model, hence we can say that stepwise regression helped us to improve our model.
\newpage



## Adding Interaction Terms
Considering the possible interactions between the variables, based on the previously calculated correlations,we added some interaction terms(aroma:flavor,flavor:acidity,aroma:acidity)in order to improve our model.We summarized the results and graphically showed the log-odds and their corresponding 95% confidence intervals.
```{r interaction model,echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
mod.cafe <- glm(Qualityclass ~ country_of_origin + aroma + flavor + acidity + category_two_defects + altitude_mean_meters + harvested + aroma:flavor + flavor:acidity + aroma:acidity, data = newdataset,
                   family = binomial(link = "logit"))
print(summary(mod.cafe)$call)
tidy(mod.cafe)
AIC(mod.cafe)
#AIC = 539
```

```{r interaction plot,echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE,out.width = '100%', fig.align = "center", ,fig.pos = "H",fig.cap="\\label{fig:interaction plot} Odds of various factors influencing the quality of coffee(model with interaction terms)"}
plot_model(mod.cafe, show.values = TRUE,
           title = "", show.p = FALSE, value.offset = 0.5)
```

In the results we can see the coefficients of aroma, flavor and acidity themselves are significantly positive, while the coefficients of all our possible interaction terms are negative,which shows that these three variables may moderate each other.After adding interaction terms, we can find that the AIC of the model decreases compared to the basic model, thus we can assume that the addition of the interaction terms improved our model.
\newpage


## GLM Stepwise After Adding Interaction Terms
In order to further improve our model, we fitted the GLM model with interaction terms using the method of stepwise regression with AIC as the criterion.
```{r interaction stepwise,echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE}
model.step <- stepAIC(glm(Qualityclass ~ country_of_origin + aroma + flavor + acidity + category_two_defects + altitude_mean_meters + harvested + aroma:flavor + flavor:acidity + aroma:acidity, data = newdataset, family = binomial(link = "logit")), direction = "both", trace = FALSE)
```

```{r interaction stepwise plot,echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE,out.width = '100%', fig.align = "center", ,fig.pos = "H",fig.cap="\\label{fig:interaction stepwise} Odds of various factors influencing the quality of coffee(stepwise regression with interaction terms)"}
print(summary(model.step)$call)
tidy(model.step)
AIC(model.step)
#AIC = 532
plot_model(model.step, show.values = TRUE,
           title = "", show.p = FALSE, value.offset = 0.50)
```

We can see from the results that the AIC decreased to the lowest among these four models we fitted. AS AIC balances simplicity and accuracy when evaluating models, we can say that after adding an interaction term and doing the stepwise regression, our fourth model is the best model. Also, the last model has the lowest BIC=720, while the other three are 766,721,776 separately, which further demonstrates the superiority of our model.

```{r level, eval=TRUE, warning=FALSE, message=FALSE}
levels(newdataset$country_of_origin)
```
\newpage


# checking assumptions
## Residuals Plots for each variables
```{r residuals, eval=TRUE, warning=FALSE, message=FALSE,out.width = '100%', fig.align = "center", ,fig.pos = "H",fig.cap="\\label{fig:iresiduals} residuals against each variables"}
res <- resid(mod.cafe)
par(mfrow=c(3,2))
plot(newdataset$aroma,res,xlab='aroma')
abline(0,0)
plot(newdataset$flavor,res,xlab='flavor')
abline(0,0)
plot(newdataset$acidity,res,xlab='acidity')
abline(0,0)
plot(newdataset$category_two_defects,res,xlab='category two defects')
abline(0,0)
plot(newdataset$altitude_mean_meters,res,xlab='altitude mean meters')
abline(0,0)
plot(newdataset$harvested,res,xlab='harvested',ylab='res')
abline(0,0)
```
We see that there is an even spread of the residuals above and below the zero line for each variables,although there are a very few outlier points, overrall their spread on the graphs are acceptable, hence our assumption that the residuals have mean zero appears valid.

# Density plot
```{r density, eval=TRUE, warning=FALSE, message=FALSE,out.width = '60%', fig.align = "center", ,fig.pos = "H",fig.cap="\\label{fig:density} density plot of residuals"}
plot(density(res))
plot(hist(res))
```
In the graph we can see that the residuals are normally distributed with the mean 0, therefore the assumption is valid.

The remaining assumptions hold naturally at the time of our modelling, thus our model appears valid.
\newpage


# Conclusion

After data cleaning and processing of non-numerical data, we fitted the data to a regression model to observe the effect of each variable in the dataset on coffee quality, and we continued to improve the model by stepwise regression and adding possible interaction terms, resulting in the model with the smallest AIC value and therefore the most profile accurate model4. Looking at the summaries and graphs of model4, we can pick out the factors that have the greatest impact: aroma and flavor are very positively influencing on the quality of coffee, with coefficients of 99.1 and 102.62. The influence of origin varies very much. When the p-value is less than a certain level of significance (0.05), which means H0 is rejected, we can see that Colombia(1.56) and Thailand(2.32) have the highest coefficient of all countries, while Uganda(-1.56), India(-3.09), Mexico(-1.03) have less coefficient, making relatively negative effects. However, many remaining origins do not seem to have obvious impact on the quality of the coffee.
